Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-05-14 draft_p2.tex
Modifications emmanuel 3/05/2019

@@

\title{Rare event algorithms for the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}

\author{Thibault Lestang}
\email{thibault.lestang@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, Insa de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\author{Freddy Bouchet}
\email{freddy.bouchet@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\author{Emmanuel Lévêque}
\email{emmanuel.leveque@ec-lyon.fr}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, Insa de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}

\begin{abstract}
This study evaluates the relevance of importance sampling techniques to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body. The main idea behind importance sampling is here to replace a long simulation by a set of shorter simulations running in parallel with dynamics that are replicated or abandoned in order to sample large events more frequently. 
%
Such techniques have already proved their efficiency for simulating extreme events in... with significant run-time savings.
Application to turbulent fluid-mechanical systems remains mainly open. 
%
The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two important sampling algorithms, namely .... 
Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort... However, limitations... pitfalls... 
Importantly, such techniques also provide access to the flow scenario yielding extreme events. In the present case, it is found that... blocking effect
\end{abstract}

@@

// Recall the previous success of importance sampling techniques with references. //
Explain why interesting to test application for turbulent systems. // To what extent a turbulent system is different from previous case studies? // 
\\

Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces. In the turbulence literature, ``extreme'' can refer to fluctuations that deviate from the mean value by $O(10)$ standard deviations.
In an engineering context, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts that may cause damages and threaten the structural integrity of embedded structures.
\\

From the viewpoint of dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium. The use of perturbative methods in identifying resonant interactions (among degrees of freedom) leading to extreme fluctuations is not successful.
In this context, simulation offers a practical means to gain physical insight into these events and estimate their frequency of occurrence, or return times. Nevertheless, this requires very long simulation since these events are also very rare. It is here our motivation to evaluate the practicality and efficiency of importance sampling techniques to sample extreme events in developed turbulent flows. 
Rare event sampling for turbulence has traditionally been approached from simplified dynamics, such as Burger's turbulence with a stochastic forcing~\cite{bec_burgers_2007}.
In this context it is possible to describe the dynamics through an action analoguous to a Gibbs measure for an equilibrium sytem.
This action can then be sampled by a Markoy chain Monte Carlo (MCMC) algorithm~\cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice}.
Even though Monte Carlo simulations are not expected to be competitive with traditional approaches such as finite difference or pseudo-spectral methods, they provide a simulation framework in which constraints can be applied so as to certain events preferentially.
A different approach is the Instanton method~\cite{grafke2015instanton}, also relying on the knowledge of an action for the dynamics.
For stochastically driven systms, in the limit of weak noise, it can be shown that the probability of extreme events is dominated by the action minima~\cite{freidlin2013random}.
The trajectory that that minimizes the action is referred to as the \textit{instanton}~\cite{gurarie_instantons_1996}.
The instanton approach consists in using suitable numerical methods to minimize the action and compute the instanton, as well as the associated probability~\cite{grafke2015instanton,chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017}.
Such approaches are however limited to stochastically driven dynamics, simple enough so that an action can be derived and sampled, or minimized. This is for instance the case of the stochastically driven 1D Burger's equation.
Although Burgers turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or geophysical flows.
In this paper, we discuss a thrid approach to the numerical simulation of rare events in turbuent flows, that aims at flows for which an action cannot be derived.
It is based on the application of \textit{rare event algorithms}, coupled with a numerical simulation of the flow dynamics.
Even though the original ideads can be traced back to the early 1950's, rare event algorithms have been of ever-growing interest voer the last 20 years. Successful application of such algorithms can be found in various domains such as chemistry~\cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016},biophysics~\cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear phsics~\cite{louvin2017}, nonlinear dynamical systems~\cite{tailleur_probing_2007}, communication networks simulation~\cite{villen-altamirano_restart:_1994}.
However, a large body of algorithms are limited to systems at equilibrium, for which the statioany ry measure can be sampled by means of Monte Carlo methods.
For systems far from equilibrium, such as turbulent flows, the sampling of rare events must be done at the level of the trajectories.
In the present paper with discuss two different algorithms suitable to out-of-equilibrum dynamics.
The Adaptive Multilevel Algorithm (2006)~\cite{cerou_adaptive_2007} inherits from older ideas about splitting approaches~\cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999}.
In recent years, it allowed the computation of rare events in problems involving a large number of degrees of freedom, such as Nolecular Dynamics simulations~\cite{aristoff_adaptive_2015}, the computation of rare trajectoris in the Allen-Cahn equations~\cite{rolland2016computing} or rare re-laminarisation in a stochastic model of wall-turbulence~\cite{rolland_extremely_2018}.
Over the last decade, the main theoritical framework for the study of rare events in statistical physics has been the one of large deviation theory~\cite{touchette_large_2009}.
Alongside, numerical methods have been developed to sample rare events based on general ideas and algorithms~\cite{moral_feynman-kac_2004}. Among them, the \ac{gktl} algorithm~\cite{giardina_direct_2006} is particularly suited to rare events in dynamical systems~\cite{giardina_simulating_2011,Laffargue_2013}.
The applicability of such algorithms to flows relevant to envorinmental or industrial applications remains open.
Recently, the \ac{gktl} algorithm allowed to perorm the cumerical simulation of extreme heat waves in a simplified model of the atmosphere~\cite{ragone_computation_2018}. Such application represent a great leap in term of complexity with respect to previous application of rare event algorithms.
In the present paper, we consider a different context of application: turbulent flows around objects.

\section{Description of the case study}
\label{sec:test_flow}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
  \caption{Our case study is a grid-generated turbulent flow impinging a squared obstacle located at the center of a channel (in two dimensions).  The vorticity is displayed with an arbitrary colormap from blue (negative values) to red (positive values). The Reynolds number based on the grid spacing is $Re=1200$.}
  \label{fig:illustr_ecoulement}
\end{figure}

The drag exerted by a grid-generated turbulent flow onto a squared obstacle is considered as our representative case study.
The flow is contained in a long channel as illustrated in Fig.~\ref{fig:illustr_ecoulement}.
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and make possible a systematic study. 
Nevertheless, this fluid-mechanical system embeds the key features that makes the application of rare event algorithms both relevant and challenging for turbulent fluid-structure interaction problems.
In particular, spatio-temporal chaos that results in the unpredictable emergence of coherent fluid structures, in turn leading to  strong fluctuations of the mechanical efforts on the structure.
In the present situation, the obstacle does not deform or move.
Through  this specific setting, our main motivation is to evaluate the operability of rare event sampling techniques to capture extreme events with a significant run-time savings.  
\\
A constant parabolic profile is imposed for the streamwise velocity at the inlet of the channel and turbulence is generated by means of a grid.
Turbulent eddies are generated in the near-wake of the grid and carried downstream.
The size of the eddies that eventually interact with the square obstacle, mounted at the center of the channel, is comparable to its size, thus resulting in strong fluctuations of the drag acting on the square. 
At the outlet of the channel, an open boundary boundary is prescribed.
In addition, the flow is progressively damped within a \textit{sponge layer}~\cite{vergnault_lattice_2012} where the viscosity is artificially increased. 
\\

%% The LBM %%

The flow dynamics are computed by the~\ac{lbm} method~\cite{kruger_lattice_2017}, using a non-orthogonal central moment relaxation scheme~\ref{De_Rosis_2016}.
The \ac{lbm} solves the Navier-Stokes equations, in a weakly compressible limit, by simulating the dynamics of fictitious particles at a mesoscopic level.
The dynamics are discretized on an isotropic lattice of spacing $\Delta x$.
The \ac{lbm} is further described in appendix~\ref{sec:lbm} and references therein.

%% LB parameters  %

The lattice spacing is set to $\Delta x = R/16$, where $R$ denotes the side of the square obstacle.
The channel has a dimension $L_x \times L_y$ with $L_y = 8R$ and $L_x = 4L_y = 32R$.
The lattice therefore has a resolution of $N_y = 129$ lattice nodes in the $y$-direction and $N_x = 513$ nodes in the $x$-direction.
The grid spacing is set to $g = 8\Delta x$.
The characteristic velocity at the inlet is set to $U_0 = 0.05 \frac{\Delta x}{\Delta t}$, where $\Delta t$ is the lattice timestep.
\textcolor{red}{@todo@Emmanuel: Briefly explain why the low Mach number and the nearly incompressible limit.}
Finally, the viscosity is set so that the Reynolds number is $Re_{grid} = U_0g/\nu = 1200$.
The Reynolds number based on the mean flow velocity and the size of the obstacle is $Re_{obs} \approx 1700$.

% Boundary conditions %

The channel walls and the surface of the obstacle are decribed by the halfway bounce-back scheme~\cite{ref_BB}\textcolor{red}{@todo@Emmanuel: Provide citation for halfway bounce-back boundary condition}.
The open boundary at the outlet is based on a second order extrapolation of the velocity along the $x$-direction.
The extrapolated velocity is then enforced using a regularization based on a finite difference estimation of the local stress tensor~\cite{latt2008straight}.
The parabolic profile at the inlet is imposed by setting populations at the corresponding equilibrium, see appendix~\ref{sec:lbm}.
\\

@@

The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle. The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and reads 
\begin{equation}
  \label{eq:drag_definition}
  f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x})  
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle. 
In our situation, the viscous stress makes a negligible contribution to the drag. This latter therefore results mostly from pressure forces, which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
As a consequence, the drag may be viewed as the pressure difference 
\begin{equation}
  \label{eq:drag_approx}
  f_d(t) = p_{fb}(t) - p_{b}(t)
\end{equation}
where $p_{fb}(t)$ and $p_{b}(t)$ denote the pressure integrated over the forebody and the base of the obstacle, respectively.

%% Definition du turnover time
The typical timescale of drag fluctuations may be estimated from dimensional analysis as
\begin{equation}
  \label{eq:turnover_time}
  \tau_0 = \frac{R}{U},
\end{equation}
where $R$ is the diameter of the cylinder and $U$ the mean-flow velocity.
Let us mention that the viscosity is here discarded since viscous stress is negligible as compared to pressure forces.

% Description du champ de vort. correspondant a une fluctuation typique
Figure~\ref{fig:typical_vorticity} displays the evolution of the vorticity field around the obstacle over a few turnover times.
The vorticity generated along the boundaries of the obstacle is advected downstream by the mean flow.
As a consequence, the base pressure varies much less than the forebody pressure, and typical drag fluctuations mostly result from the fluctuating pressure at the front of the obstacle.

In the section~\ref{sec:direct_sampling}, we highlight a very different phenomenology for \textit{extreme} fluctuations of the drag acting on the obstacle.
In particular, we show that extreme fluctuations mainly result from fluctuations of the base pressure, whereas forebody pressure fluctuations play a lesser role.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
  \caption{Vorticity dynamics corresponding to typical drag fluctuations over four turnover times~\eqref{eq:turnover_time}. Negative vorticity is generated along the bottom boundary, and advected away from the the obstacle by the mean flow.}
  \label{fig:typical_vorticity}
\end{figure}

\subsection{The drag random process: Probability Density Function, time correlations}
\label{sec:pdfs}

% Description du signal de trainee typique
Fig.~\ref{fig:typical_drag_signal} illustrates the time evolution of the drag acting on the square obstacle.
The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
It is therefore natural to model the drag as a scalar random process $f_d(t)$.
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
  \caption{Temporal evolution of the drag acting on the square under the action of the incident turbulent flow. The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
  \label{fig:typical_drag_signal}
\end{figure}

Drag fluctuations have been sampled along a long simulation of duration $T_{tot} = 4\times 10^6~\tau_0$, which will be referred to as the \textit{control run} in the following.
The corresponding estimate of the \ac{pdf} of drag fluctuations is shown in figure~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  ${\cal P}(f_d) = \lambda^{-1}e^{-\lambda f_d}$.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface equivalent to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails, indicating that the asymmetry of the \ac{pdf} and the development of positive exponential tail are closely related to the no-slip condition on the obstacle.
\begin{figure}
  \centering
    \subfloat[\ac{pdf} of drag fluctuations around mean.]
    {\label{fig:pdf_drag_a}
    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
    \subfloat[Autocorrelation of drag.]
    {\label{fig:pdf_drag_b}
    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
      \caption{\textbf{(a)} Estimate of the zero-mean \ac{pdf} for drag fluctuations $f'_d = (f_d - \bar{f_d})$ where $\bar{f_d}$ denotes the averaged value. The drag is evaluated both in the presence and absence of the obstacle. Note that both \acsp{pdf} have different variance, and are here not normalized.
  \textbf{(b)} Autocorrelation function of the drag defined as $\left( \langle f_d(t+\tau)f_d(t) \rangle-\langle f_d \rangle ^2 \right) /\sigma^2$. The lag $\tau$ is normalized by the turnover-time~\eqref{eq:turnover_time}.}
  \label{fig:pdf_drag}
\end{figure}
%
Lastly, the autocorrelation function of the drag is shown in Fig.~\ref{fig:pdf_drag_b}. It is found drag fluctuations are correlated over a time scale of the order of $\tau_0$, suggesting that the drag signal looses its memory typically after the sweeping of a few eddies past the obstacle.
%
This observation has important consequences for the application of rare event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the following we define the \textit{correlation time} $\tau_c = 4\tau_0$.
Note that the ratio $\tau_0 / \tau_c$ may be viewed as a \textit{Strouhal number} $St$.
A value $St=0.25$ is consistent with commonly observed values for flows past blunt structures.

@@

In this section we describe the phenomenology of extreme fluctuations of the drag that acts on the square cylinder mounted in the flow described in section~\ref{sec:test_flow}.

>>>>

\newcommand{\todo}[1]{{\color{red}{#1}}}

@@

\title{Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body}

\author{Thibault Lestang}
\email{thibault.lestang@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\author{Freddy Bouchet}
\email{freddy.bouchet@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\author{Emmanuel Lévêque}
\email{emmanuel.leveque@ec-lyon.fr}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}

\begin{abstract}
This study evaluates the relevance of  rare-event sampling to optimize the numerical simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body. The main idea is here to replace a long simulation by a set of much shorter simulations (running in parallel) with dynamics that are replicated or abandoned in order to sample large-amplitude events more frequently. 
%
Such technique has already proved its efficiency for simulating extreme events in {<...>} with significant run-time savings.
Application to fluid-structure interaction problems remains mainly open. 
%
The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major sampling algorithms, namely {<...>}. 
Practical evidence is given that these algorithms allow for the correct sampling of rare large-amplitude drag events and the estimation of return times with a reduced computational effort. However, {<limitations and pitfalls>}. 
% Importantly, such techniques also provide direct access to the flow scenario yielding extreme events. In the present case, it is found that {<blocking effect>}.
\end{abstract}

@@

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}. ``Extreme'' refers here to fluctuations that can deviate from mean value by ${\cal{O}}(10)$ standard deviations. In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts. Such anomalous efforts can threaten the structural integrity of embedded structures.
%
From the viewpoint of (chaotic) dynamical systems, turbulence in fluids has to do with non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}. The use of perturbative methods in identifying resonant interactions (among degrees of freedom) possibly responsible for extreme fluctuations is not successful. Simulation offers a practical means to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence. Nevertheless, this requires very long simulation since these events are also very rare. It is here our motivation to evaluate the practicality and efficiency of rare-event sampling techniques to alleviate this prohibitive computational cost. 
%

% general def on rare-event sampling %
%
Rare-event sampling is an umbrella term for numerical methods intended to sample preferentially regions of the phase space which would otherwise be rarely visited  through brute-force simulation.
%

%
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
% this paragraph would certainly need to be reshaped!
%
In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}. In that case, dynamics may be accounted for through an action that can be identified as analogous to a Gibbs measure for a system at equilibrium. This latter can be sampled by using a Markov chain Monte-Carlo  algorithm \cite{duben_monte_2008,mesterhazy_anomalous_2011,mesterhazy_lattice_2013}, which provides a framework for rare-event sampling.
%
An alternative approach is based on \emph{instanton} \cite{grafke_instanton_2015}. For stochastically driven systems in the limit of weak noise, it can be proved that the probability of extreme events is dominated by the minima of the action \cite{freidlin2013random} -- the trajectories that achieve such minima being referred to as instantons \cite{gurarie_instantons_1996}. Therefore, suitable numerical schemes can  be used to minimize the action and evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grafke_relevance_2015,grigorio_instantons_2017}.
%
An obvious drawback of the aforementioned approaches is their limitation to stochastically driven dynamics which, in addition, must be simple enough so that an action may be derived and sampled, or minimized. 
%
%
%Whilst Burgers' turbulence provides an interesting model framework for the investigation of the properties of small scale turbulence, it is of little relevance to more realistic situations, such as industrial or environmental flows.
%
In the present paper, a more straightforward approach is investigated. It relies on sampling algorithms connected to the simulation of the actual dynamics itself, \textit{i.e.} applying directly to the trajectories in the dynamic space.
% dedicated to dynamical systems for which an action cannot be easily derived or minimized,
%    
%
Even though such ideas trace back to the early 1950's, they have received ever-growing interest over the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
%
% already mentioned in the beginning? This sentence is misleading.
%
%However, a large body of algorithms are limited to systems at equilibrium, for which the stationary measure can be sampled by means of Monte Carlo methods. 
%
An original contribution of our study is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
Let us remark that most of previous successful applications involve dynamical systems at statistical equilibrium and low-dimensional attractors. \todo{TRUE?}
% 
Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
%

The Adaptive Multilevel Splitting (AMS) algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a more detailed description will be given later in the paper.  
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom, such as molecular dynamics simulations \cite{aristoff_adaptive_2015}, the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016} or rare re-laminarization in a stochastic model of wall-turbulence \cite{rolland_extremely_2018}.

Over the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
Among them, the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
%
Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modeling of the atmosphere \cite{ragone_computation_2018}. 
This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.

\section{Description of the numerical case study}
\label{sec:test_flow}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{illustr_ecoulement/illustr_ecoulement}
\caption{Our case study is a grid-generated turbulent flow impinging onto a fixed squared obstacle (of size $R$) located at the center of a channel in two dimensions. The flow is artificially damped near the end of the channel. In the developed flow, turbulent eddies have typically the size of the square. The vorticity is displayed with an arbitrary color map from blue (negative values) to red (positive values).}
\label{fig:illustr_ecoulement}
\end{figure}

% introduce the flow 
%
The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
% why this flow?
%
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study. 
% 
We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
%
%
% rapid description of the flow
Turbulent eddies generated in the near-wake of the grid are carried downstream. They interact with each other and grow in size as expected in two-dimensional turbulence. 
The dimension of the grid is such that the size of the eddies that eventually hit the square is comparable to the size of this latter, which results in strong fluctuations of the drag acting on the square. 
%
%Finally, the obstacle does not deform or move either. 
%
%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
%

% lattice Boltzmann method 
%
The flow dynamics is integrated by the lattice Boltzmann (LB) method in our numerical simulations. 
While traditional methods in \emph{computational fluid dynamics} rely on a discretization of the Navier-Stokes equations, the LB method considers the fluid at a mesoscopic level. Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
%namely the Navier-Stokes dynamics in the weakly compressible limit
Further details are given in Appendix and references therein.
In our situation, the LB method has been chosen principally for its outstanding computational efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $512 \times 128$ mesh points. The squared obstacle has size $R=16$ (in mesh unit) and is located at the center of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel by using an halfway bounce-back procedure \cite{lbm_book}.  
%
Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
The centerline velocity is $0.05$ in lattice unit, \textit{i.e.} normalized by $(\Delta x/\Delta t)$ with $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The distributions are imposed at equilibrium (see Appendix). 
In the bulk, the viscosity is adjusted to generate grid turbulence with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method. 
Near the end of the channel, the flow is progressively damped within a sponge layer where the viscosity is artificially enhanced. Finally, the oulet boundary condition relies on a second-order extrapolation of the velocity and mass density. Then, the extrapolated distributions are evaluated through a regularization procedure (development onto the first two tensor Hermite polynomials) relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will be identified later as the correlation time of the drag signal.
	The vorticity is given in lattice units. \todo{THE COLORMAP SEEMS TO BE REVERSED IN THE WHOLE PAPER: BLUE SHOULD BE NEGATIVE AND RED POSITIVE: $\omega = \partial_x u_y - \partial_y u_x$}}
	\label{fig:typical_vorticity}
\end{figure}

% drag signal
%
The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle. The \textit{drag} is defined as the resulting force in the streamwise $x$-direction and expresses formally as 
\begin{equation}
\label{eq:drag_definition}
f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x})  
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
Here, the viscous stress makes a negligible contribution to the drag, which therefore results mostly from pressure forces.
% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
As a consequence, the drag can be expressed as the pressure difference 
\begin{equation}
\label{eq:drag_approx}
f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
between the pressure integrated over the forebody ($p_{fb}(t)$) and the base of the obstacle ($p_{base}(t)$).

% Definition of turnover time
%
The typical timescale (turnover time) of drag fluctuations may be estimated from dimensional analysis as
\begin{equation}
\label{eq:turnover_time}
\tau_0 = \frac{R}{U}
\end{equation}
where $R$ is the size of the square and $U$ the averaged velocity. 
The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
%
% typical fluctuation of the drag
%
Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times. It is observed that some vorticity is generated near the front part of the obstacle, rapidly detach from it and is advected downstream by the flow.
Therefore, the base pressure varies much less than the forebody pressure, and typical drag fluctuations  result mostly from the variation of the pressure at the front of the obstacle.
In section~\ref{sec:direct_sampling}, we shall highlight that the situations leading to  {extreme} fluctuations of the drag are different.
In particular, extreme fluctuations mainly result from large variations of the base pressure, whereas forebody pressure variations play a lesser role.

\subsection{The drag as a random process: Probability Density Function, correlation in time}
\label{sec:pdfs}

\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{typical_drag_signal/typical_drag_signal.png}
	\caption{Temporal evolution of the drag (in lattice units) acting on the square under the action of the impinging turbulent flow. The time is normalized by the turnover time related the mean-flow velocity and the size of the obstacle: $\tau_0=R/U$.}
	\label{fig:typical_drag_signal}
\end{figure}

% drag signal
%
Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square over five hundred turnover times. The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
It is therefore natural to model the drag as a (scalar) random process $f_d(t)$.

\begin{figure}
  \centering
    \subfloat[\ac{pdf} of (zero-mean) drag fluctuations]
    {\label{fig:pdf_drag_a}
    	\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
    \subfloat[Autocorrelation of drag]
    {\label{fig:pdf_drag_b}
    	\includegraphics[width=.45\linewidth]{./autocorr_drag/autocorr_drag.png}}  
      \caption{\textbf{(a)} \ac{pdf} of (zero-mean) drag fluctuations $\tilde f_d \equiv f_d - \bar{f_d}$, where $\bar{f_d}$ denotes the time averaged value. The drag is evaluated both in the presence (red) and in the absence of the obstacle (blue). Note that the \acsp{pdf} not been normalized.
  \textbf{(b)} Autocorrelation function of the drag defined as $\overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The time lag has been normalized by $\tau_c \simeq 4\tau_0$ so that $C(1)=0$: $\tau_c$ can be considered as the correlation time of the drag signal.}
  \label{fig:pdf_drag}
\end{figure}

% drag statistics
% PDF and correlation time
%
Drag fluctuations have been sampled along a simulation of duration $T_{tot} = 4 \times 10^6~\tau_0$. This very long simulation will be referred to as the \textit{control run} in the remainder. It has been made possible by the relative simplicity of the investigated two-dimensional flow; its execution time is about a few weeks (wall-clock time).
The \ac{pdf} of the drag fluctuations is displayed in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and exhibits an exponential tail for large positive values, \textit{i.e.}  ${\mathbb{P}}(\tilde f_d) \propto e^{-\lambda \tilde f_d}$.
%
Fig.~\ref{fig:pdf_drag_a} also shows the \ac{pdf} of drag fluctuations acting on a (virtual) control surface along the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This suggests that the asymmetry of the \ac{pdf} and the positive exponential tail are closely related to the no-slip condition on the obstacle.
%
%
Lastly, the autocorrelation function of the drag is shown in Fig.~\ref{fig:pdf_drag_b}. It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$. The drag therefore looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
%
This observation has important consequences for the application of rare-event algorithms, as will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the remainder, $\tau_c$ will be referred to as the \textit{correlation time} of the drag. The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}; $St=0.25$ is consistent with commonly observed values for flows past blunt structures.

@@

\begin{figure}
	\centering
	\includegraphics[width=.6\linewidth]{return_time/return_time.png}
	\caption{Amplitude of drag fluctuation (in standard-deviation unit) as a function the corresponding return time. The return time $r(a)$ is defined as the typical period of occurrence of a fluctuation of amplitude larger than $a$.}
	\label{fig:return_time_instant}
\end{figure}
